### 2020-11-12 thesis/appendix_3.tex
half of comment 45

\chapter*{\textbf{Supplementary analysis of land use change in Monteregie \\ \hspace{1em}}}
\addcontentsline{toc}{chapter}{Appendix 4: Supplementary analysis of land use change in Monteregie}

@@

>>>>

\chapter*{\textbf{Appendix 3: Supplementary analysis of land use change in Monteregie \\ \hspace{1em}}}
\addcontentsline{toc}{chapter}{Appendix 3: Supplementary analysis of land use change in Monteregie}

@@

The clustering and subsequent ordination of the land use change matrices of the 177 municipalities revealed that Montérégie has 5 main profiles and land compositions (see figures \ref{fig:PCAvals}, and \ref{fig:mapvals}):
\renewcommand{\labelitemi}{$\textendash$}
\begin{itemize}[leftmargin=0.5cm]
 \item \textbf{Forest - Dominant}: have the lowest level of fragmentation and are dominated by forest.
 \item \textbf{Forest - Agriculture}: they still have a significant area of forest but fragmentation is much more pronounced.
 \item \textbf{Agriculture - Dominant}: forested habitat is scarce and most of the remaining forest is classified as “Trees” in the AAFC dataset (forest fragment of less than 1 hectare).
 \item \textbf{Urban - Medium density}: correspond to the front of the wave of urban sprawling.
 \item \textbf{Urban - High density}: urban cores make up the majority of the municipality’ area (> 50\%).
\end{itemize}

The clustering and subsequent ordination of land use change profiles showed that Montérégie has 4 different transition profiles (see figures \ref{fig:PCAtrans}, and \ref{fig:maptrans}):
\begin{itemize}[leftmargin=0.5cm]
 \item \textbf{Urban Spread / Deforestation}: forest fragmentation is progressing mainly via the growth of urban land (in the west) or development for holiday cottages (“villegiatives pressures”, in the east).
 \item \textbf{Urban Spread / Agricultural loss}: agriculture is losing ground to urban land.
 \item \textbf{Agricultural Expansion / Fragmentation}: forest is being lost to agriculture in those municipalities where forest is still quite present.
 \item \textbf{Agricultural Expansion / Deforestation}: forest is already scarce and is being replaced by agricultural land.
These results show marked regional trends with a front line of fragmentation and deforestation on both sides of the region and along the Richelieu river. \\ % (Fig. \ref{fig:map}).
\end{itemize}

### 2020-11-12 thesis/chap_1.tex
half of comment 45

% [back to what we want to do]
Climate change can interact with land use change in ways that impact biodiversity \citep{oliver_interactions_2014} It is therefore likely that connectivity models integrating these two drivers would improve connectivity conservation methods such as CPAN design. Yet, in a recent review, \cite{costanza_landscape_2019} have found that connectivity methodological frameworks rarely combine  land use change and climate change. Here, we build on recent connectivity modelling work using circuit theory, by pairing it with a hybrid land use and climate change model \citep{albert_applying_2017}. The model we develop uses a combination of statistical modelling and simulations based on a Markov Chain Monte Carlo (MCMC) method, combining Random forests (RF) models with a Cellular automaton (CA) model. We use this RF-CA model to project future land use changes and estimate the changes in potential functional connectivity for five different umbrella species, under different climate change scenarios. We also derive potential insights for the design of a hypothetical regional network of connected protected areas resilient to future landscape change. In addition, we provide an analysis of past land use change trends in our study region, the Montérégie region in southern Québec.

@@

@@

The AAFC data does not differentiate between different forest types. However, our connectivity analysis relies on an analysis of habitat suitability, which is reliant on knowing the preferences of different species for different forest types (see section of habitat suitability analysis for more details). In order to integrate dynamics of forest types into our simulations, we relied on the results of a separate set of simulations that were run in LANDIS, a free and open source software designed to model forest growth. Those simulations were run by Boulanger and Larocque (unpublished, see Appendix 3) for the extent of the Saint-Lawrence lowlands ecoregion, in the context of a research contract with the MELCC (Ministère de l'Environnement et de la Lutte contre les changements climatiques). The LANDIS simulations do not include land use change and therefore can not be directly merged into our simulations. Instead, we parameterized the changes in forest types from the results of the LANDIS simulations. A more detailed methodological description of the LANDIS simulations is to be found in the Appendix 3.

@@

\subsubsection{Preliminary analysis of past land use change}

To better understand land use patterns in the region, we ran some preliminary analysis on land use change data at the municipality level. Montérégie is made up of 177 different municipalities. We generated land use matrices (amount of each land use category) and land cover transition matrices for the first timestep of our dataset (1990 to 2000) for each municipality and for the land use categories corresponding to our model. We then performed multivariate analyses by running a first ordination using PCA (Principal Component Analysis), and then using Ward clustering \citep{legendre_ordination_2012}.
Because municipalities in Montérégie have variable size, the amount of change in each municipality needs to be standardized. The data was normalized prior to the ordination to reflect relative amounts of change in each municipality. We used the \verb|vegan| package in R to produce such plots.\\

@@

We first describe the results of the preliminary analysis on land use change dynamics, then go on to describe the results of the hybrid land use change and connectivity model.\\

\subsection{Analysis of past land use changes in Montérégie}

The clustering and subsequent ordination of the land use change matrices of the 177 municipalities revealed that Montérégie has 5 main profiles and land compositions (see figures \ref{fig:PCAvals}, and \ref{fig:mapvals}):
\renewcommand{\labelitemi}{$\textendash$}
\begin{itemize}[leftmargin=0.5cm]
 \item \textbf{Forest - Dominant}: have the lowest level of fragmentation and are dominated by forest.
 \item \textbf{Forest - Agriculture}: they still have a significant area of forest but fragmentation is much more pronounced.
 \item \textbf{Agriculture - Dominant}: forested habitat is scarce and most of the remaining forest is classified as “Trees” in the AAFC dataset (forest fragment of less than 1 hectare).
 \item \textbf{Urban - Medium density}: correspond to the front of the wave of urban sprawling.
 \item \textbf{Urban - High density}: urban cores make up the majority of the municipality’ area (> 50\%).
\end{itemize}

The clustering and subsequent ordination of land use change profiles showed that Montérégie has 4 different transition profiles (see figures \ref{fig:PCAtrans}, and \ref{fig:maptrans}):
\begin{itemize}[leftmargin=0.5cm]
 \item \textbf{Urban Spread / Deforestation}: forest fragmentation is progressing mainly via the growth of urban land (in the west) or development for holiday cottages (“villegiatives pressures”, in the east).
 \item \textbf{Urban Spread / Agricultural loss}: agriculture is losing ground to urban land.
 \item \textbf{Agricultural Expansion / Fragmentation}: forest is being lost to agriculture in those municipalities where forest is still quite present.
 \item \textbf{Agricultural Expansion / Deforestation}: forest is already scarce and is being replaced by agricultural land.
These results show marked regional trends with a front line of fragmentation and deforestation on both sides of the region and along the Richelieu river. \\ % (Fig. \ref{fig:map}).
\end{itemize}

@@

% Figures: values

% Clustering moved to appendix

% PCA
\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=0.9\textwidth]{thesis/figures/PCA_data_profiles.png}
}
\caption{Ordination of land use data (proportions) for municipalities.}
\label{fig:PCAvals}

% MAP
\makebox[\textwidth]{
    \includegraphics[width=0.9\textwidth]{thesis/figures/profiles_land_use.png}
}
\caption{Geographical distribution of the 5 profiles identified in figure \ref{fig:PCAvals}.}
\label{fig:mapvals}
\end{figure}

% Figures: Transitions

% PCA
\begin{figure}[h!]
\makebox[\textwidth]{
  \includegraphics[width=0.9\textwidth]{thesis/figures/PCA_trans_profiles.png}
}
\caption{Ordination of land use transition data for municipalities.}
\label{fig:PCAtrans}

%MAP
\makebox[\textwidth]{
    \includegraphics[width=0.9\textwidth]{thesis/figures/transition_prof_map.png}
}
\caption{Geographical distribution of the 4 change profiles identified in figure \ref{fig:PCAtrans}.}
\label{fig:maptrans}
\end{figure}

\clearpage

%---------------------------------------------------------------------------------------------------------------------------------------------------
% Model tables (roc curves moved to the appendix)

>>>>

% [back to what we want to do]
Climate change can interact with land use change in ways that impact biodiversity \citep{oliver_interactions_2014} It is therefore likely that connectivity models integrating these two drivers would improve connectivity conservation methods such as CPAN design. Yet, in a recent review, \cite{costanza_landscape_2019} have found that connectivity methodological frameworks rarely combine  land use change and climate change. Here, we build on recent connectivity modelling work using circuit theory, by pairing it with a hybrid land use and climate change model \citep{albert_applying_2017}. The model we develop uses a combination of statistical modelling and simulations based on a Markov Chain Monte Carlo (MCMC) method, combining Random forests (RF) models with a Cellular automaton (CA) model. We use this RF-CA model to project future land use changes and estimate the changes in potential functional connectivity for five different umbrella species, under different climate change scenarios. We also derive potential insights for the design of a hypothetical regional network of connected protected areas resilient to future landscape change. In addition, we provide an analysis of past land use change trends in our study region (appendix 3), the Montérégie region in southern Québec.

@@

\subsubsection*{\textit{Preliminary analysis of past land use change}}

To get a better sense of land use patterns in the region, we ran some preliminary analysis on land use change data at the municipality level. These results are not presented in the main text but are to be found in Appendix 3. 
Montérégie is made up of 177 different municipalities. We generated land use matrices (amount of each land use category) and land cover transition matrices for the first timestep of our dataset (1990 to 2000) for each municipality and for the land use categories corresponding to our model. We then performed multivariate analyses by running a first ordination using PCA (Principal Component Analysis), and then using Ward clustering \citep{legendre_ordination_2012}. The results of this analysis
Because municipalities in Montérégie have variable size, the amount of change in each municipality needed to be standardized. The data was normalized prior to the ordination to reflect relative amounts of change in each municipality. We used the \verb|vegan| package in R to produce such plots.\\

@@

The AAFC data does not differentiate between different forest types. However, our connectivity analysis relies on an analysis of habitat suitability, which is reliant on knowing the preferences of different species for different forest types (see section of habitat suitability analysis for more details). In order to integrate dynamics of forest types into our simulations, we relied on the results of a separate set of simulations that were run in LANDIS, a free and open source software designed to model forest growth. Those simulations were run by Boulanger and Larocque (unpublished, see Appendix 4) for the extent of the Saint-Lawrence lowlands ecoregion, in the context of a research contract with the MELCC (Ministère de l'Environnement et de la Lutte contre les changements climatiques). The LANDIS simulations do not include land use change and therefore can not be directly merged into our simulations. Instead, we parameterized the changes in forest types from the results of the LANDIS simulations. A more detailed methodological description of the LANDIS simulations is to be found in the Appendix 4.

@@

@@

@@

% Model tables (roc curves moved to the appendix)
